---
title: "Shared confounding Simulation"
format: html
editor: source
editor: source
---

```{r}
library(cmdstanr)
library(tidyverse)
library(mgcv)
library(randomForest) 
# library(AER)
library(fields)
library(MASS)
library(pci2s)
library(gsynth)
library(FNN)
library(nFactors)
library(mvtnorm)

library(rstiefel)
# library(psych)    # ML factor analysis (or: factanal)
library(expm)     # for sqrtm
library(scales)

source("utilities.R")
options(mc.cores=parallel::detectCores())

dreg <- function(x,d){randomForest(x, d)} 
yreg <- function(x,y){randomForest(x, y)}
```

## Spatially invariant confounding

```{r}
fit_spatial_invariant <- function(model, s, d, y, m, cache = NULL) {
  n <- nrow(s); t <- nrow(d)

  switch(model,
         single_gp = {
           d1 <- as.vector(d[1,])
           y1 <- as.vector(y[1,])
           data_list <- list(D=ncol(s), N=n, s=s, t=d1, y=y1)
           gp_stan <- cmdstanr::cmdstan_model("stan/outcome_gp.stan")
           single_gp_mle <- gp_stan$optimize(data=data_list, init=2, iter=1000, show_messages=FALSE)
           return(single_gp_mle$mle("beta"))
         },

         single_dml = {
           dml_rf <- dml_plm(s, d[1,], y[1,], dreg, yreg, nfold=5)
           res_model <- lm(dml_rf$resY ~ dml_rf$resD)
           return(coef(res_model)[2])
           # dml_rf$coef.est
         },
         
         single_splines = {
           s1 <- s[, 1]
           s2 <- s[, 2]
           plm_model <- mgcv::gam(y[1,] ~ d[1,] + s(s1, s2, k=min(n, 200)))
           return(coef(plm_model)[2])
         },
         
         multi_lm = {
           lm_model <- lm(y ~ d)
           return(coef(lm_model)[2, 1])
         },

         multi_spline = {
           s1 <- rep(s[, 1], each = t)
           s2 <- rep(s[, 2], each = t)
           d_res <- matrix(residuals(mgcv::gam(as.vector(d) ~ s(s1, s2, k=min(n, 200)))), nrow=t)
           y_res <- matrix(residuals(mgcv::gam(as.vector(y) ~ s(s1, s2, k=min(n, 200)))), nrow=t)
           lm_model <- lm(y_res ~ d_res)
           return(coef(lm_model)[2, 1])
         },
         
         multi_dml = {
           resY <- matrix(cache$resY, nrow=t)
           resD <- matrix(cache$resD, nrow=t)
           lm_model <- lm(resY ~ resD)
           return(coef(lm_model)[2, 1])
         },
         
         stack_gp = {
           sm_nc_gp <- cmdstanr::cmdstan_model("stan/multi_nc_partial_gp.stan")
           data_list <- list(D=ncol(s), N=n, T0=t, s=s, d=d, y=y)
           opt_nc_gp <- sm_nc_gp$optimize(data=data_list, init=2, iter=1000, show_messages = FALSE)
           return(opt_nc_gp$mle("beta"))
         },
         
         stack_dml = {
           lm_model <- lm(cache$resY ~ cache$resD)
           return(coef(lm_model)[2])
         },

         stack_spline = {
           s1 <- rep(s[, 1], each = t)
           s2 <- rep(s[, 2], each = t)
           plm_model <- mgcv::gam(as.vector(y) ~ as.vector(d) + s(s1, s2, k=min(n, 200)))
           return(coef(plm_model)[2])
         },

         nc_mfactor = {
           sm_nc_factor <- if (!is.null(cache$sm_nc_factor)) {
             cache$sm_nc_factor
           } else {
             cmdstanr::cmdstan_model("stan/multi_nc_linear_factor.stan")
           }
           data_list <- list(K=n, Q=n, N=t, M=m, d=d, y=y)
           opt_result <- sm_nc_factor$optimize(data=data_list, init=2, iter=1000, show_messages=FALSE)
           return(opt_result$mle("beta"))
         },

         nc_2mfactor = {
           sm_nc_factor <- cmdstanr::cmdstan_model("stan/multi_nc_linear_factor.stan")
           data_list <- list(K=n, Q=n, N=t, M=m*2, d=d, y=y)
           opt_result <- sm_nc_factor$optimize(data=data_list, init=2, iter=1000, show_messages=FALSE)
           return(opt_result$mle("beta"))
         },
 
         spline_mfactor = {
           s1 <- rep(s[, 1], each = t)
           s2 <- rep(s[, 2], each = t)
           # time = rep(1:t, times = n)
           d_res <- matrix(residuals(mgcv::gam(as.vector(d) ~ s(s1, s2, k=min(n, 200)))), nrow=t)
           y_res <- matrix(residuals(mgcv::gam(as.vector(y) ~ s(s1, s2, k=min(n, 200)))), nrow=t)
           sm_nc_factor <- cmdstanr::cmdstan_model("stan/multi_nc_linear_factor.stan")
           data_list <- list(K=n, Q=n, N=t, M=m, d=d_res, y=y_res)
           opt_result <- sm_nc_factor$optimize(data=data_list, init=2, iter=1000, show_messages=FALSE)
           return(opt_result$mle("beta"))
         },

         dml_mfactor = {
           resY <- matrix(cache$resY, nrow=t)
           resD <- matrix(cache$resD, nrow=t)
           sm_nc_factor <- cmdstanr::cmdstan_model("stan/multi_nc_linear_factor.stan")
           data_list <- list(K=n, Q=n, N=t, M=m, d=resD, y=resY)
           opt_result <- sm_nc_factor$optimize(data=data_list, init=2, iter=1000, show_messages=FALSE)
           return(opt_result$mle("beta"))
         },
         
         ife = {
           df <- data.frame(unit = rep(1:n, each = t),
                            time = rep(1:t,  n),
                            Y = c(y),
                            D = c(d))
           gs <- NULL
           invisible(capture.output({
             gs <- interFE(Y ~ D, data = df, index = c("unit","time"), r = m*2,
                           force = "none", se = FALSE)
           }))
           # ife_res <- ife_bai(y, d, 3)
           return(gs$beta)
         },
         
         dml_ife = {
           df <- data.frame(unit = rep(1:n, each = t),
                            time = rep(1:t,  n),
                            Y = c(cache$resY),
                            D = c(cache$resD))
           gs <- NULL
           invisible(capture.output({
             gs <- interFE(Y ~ D, data = df, index = c("unit","time"), r = m,
                           force = "none", se = FALSE)
           }))
           # ife_res <- ife_bai(y, d, 3)
           return(gs$beta)
         },
         
         proximal_dnc = {                 
           resY <- matrix(cache$resY, nrow = t)   # T × n
           resD <- matrix(cache$resD, nrow = t)
        
           if (n < m + 1)
             stop("Need at least m+1 units to supply ", m, " NC pairs")
        
           W <- resY[, 2]   # negative-control outcomes
           Z <- resD[, 3]   # negative-control exposure
           Y <- resY[, 1]   # focal outcome
           A <- resD[, 1]   # focal exposure

           p2sls_result <- p2sls.lm(
             Y = Y,
             A = A,
             X = NULL,
             W = W,
             Z = Z,
             nco_type = "linear",
             nco_args = list(list(offset = rep(0, length(Y))))
           )
           return(p2sls_result$ESTIMATE[2])
          },

         NA)
}
```

### Spatial variation

```{r}
set.seed(2025) 
results_sp_loc <- list()
n_simulations = 100

N <- 50
t <- 100
m <- 3
sigmaU <- diag(m)  # Update sigmaU based on N

A <- matrix(rnorm(N * m, mean = 0, sd=0.5), nrow = N, ncol = m)
R <- qr.Q(qr(matrix(rnorm(m*m), m))) 
gamma <- 0.3 * (A %*% R) + 0.5 * A
B_true <- diag(N)

for (i in 1:n_simulations) {
  # Generate data for the current setting
  s <- matrix(runif(2 * N, -1, 1), ncol = 2) # N by 2
  fs <- sin(2*pi*s[,1]*s[,2]) + s[,1] + s[,2] # N by 1
  u <- mvtnorm::rmvnorm(t, sigma = sigmaU) # t * m

  d <- u %*% t(A) + matrix(rep(fs^2/2,t), nrow=t, byrow=TRUE) + rnorm(t * N, sd = 0.2) # t by N
  y <- d %*% t(B_true) + u %*% t(gamma) + matrix(rep(fs,t), nrow=t, byrow=TRUE) + rnorm(t * N, sd = 0.5) 
  
  # fit DML model
  x <- data.frame(s1=rep(s[, 1], each = t), s2=rep(s[, 2], each = t))
  # Polynomial spatial terms
  x$s1_sq  <- x$s1^2
  x$s2_sq  <- x$s2^2
  x$s1_s2  <- x$s1 * x$s2
  x$s1_cu  <- x$s1^3
  x$s2_cu  <- x$s2^3
  dml_rf <- dml_plm(as.matrix(x), as.vector(d), as.vector(y), dreg, yreg, nfold=5)

  # Loop over models
  for (model in c("single_dml", "multi_dml", "stack_dml", "dml_mfactor",
                  "nc_mfactor", "nc_2mfactor",
                  "ife", "dml_ife", "proximal_dnc")) {
    estimate <- fit_spatial_invariant(model, s, d, y, m, dml_rf)  # Fit the model 
    print(model)
    results_sp_loc[[length(results_sp_loc) + 1]] <- data.frame(
      simulation = i,
      model = model,
      estimate = estimate
    )
  }
}


# Combine results into a data frame
results1_df <- do.call(rbind, results_sp_loc)
results1_df$error <- (results1_df$estimate - 1) / 1

results1_df <- results1_df %>%
  mutate(model = case_when(
    model == "single_dml" ~ "Single DML",
    model == "multi_dml" ~ "Multi DML",
    model == "stack_dml" ~ "Stacked DML",
    model == "nc_mfactor" ~ "FC(3)",
    model == "nc_2mfactor" ~ "FC(6)",
    model == "dml_mfactor" ~ "FC(3)+DML",
    model == "ife" ~ "IFE(6)",
    model == "dml_ife" ~ "IFE(3)+DML",
    model == "proximal_dnc" ~ "PCI",
    TRUE ~ model
  ))

# saveRDS(results1_df, file = "sim_linear_factor_N50T100_sp.rds")
```

```{r}
# Relative error
## results1_df <- readRDS("data/sim_linear_factor_N50T100_sp.rds")
p1 <- ggplot(results1_df |> filter(model != "PCI"), aes(x = model, y = error)) +
  geom_boxplot(fill = "dark gray", outliers=FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red", size = 0.5) +
  # facet_wrap(~setting, scales = "free") +
  labs(title = "Fixed Spatial Effects and Spatially Invariant Confounding", 
       x = "Model", 
       y = "Relative Error",
       fill = "Model") +
  theme_bw(base_size=16) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") 
p1

# ggsave("sim_linear_factor_N50T100_sp_error.pdf", p1, width = 6, height = 5, units = "in", dpi = 300)
```

```{r}
u_s <- tibble(t = ds[,1], s1 = s[,1], s2 = s[,2])
u_s %>%
  ggplot(aes(x=s1, y=s2, color=t)) + 
  geom_point(alpha=1, size=2.5) +
  theme_bw() + colorspace::scale_color_continuous_diverging("Blue-Red 2")
```

### Spatiotemporal variation

```{r, message=FALSE, warning=FALSE}}
rho <- 0.9
results_st_loc <- list()

set.seed(2025) 
n_simulations = 100

N <- 50
t <- 100
m <- 3
sigmaU <- diag(m) 
shared_basis <- matrix(rnorm(N * m, mean = 0, sd=0.2), nrow = N, ncol = m)
A <- 2 * shared_basis
R <- qr.Q(qr(matrix(rnorm(m*m), m))) 
gamma <- 0.4 * (shared_basis %*% R) + 0.6 * A
B_true <- diag(N)

for (j in 1:n_simulations) {
  s <- matrix(runif(2 * N, -1, 1), ncol = 2) # N by 2
  ds <- matrix(0, nrow=N, ncol=t) 
  ds[,1] <- gp_sample(s, a2 = 0.6, l = 0.5, sigma2 = 1e-5)
  for(i in 2:t) {
    e_i <- gp_sample(s, a2 = 0.1, l = 0.5, sigma2 = 1e-5)
    ds[,i] <- rho * ds[,i-1] + e_i
  }
  ys <- matrix(0, nrow=N, ncol=t) # n by t 
  ys[,1] <- gp_sample(s, a2 = 0.6, l = 0.5, sigma2 = 1e-5)
  for(i in 2:t) {
    e_i <- gp_sample(s, a2 = 0.1, l = 0.5, sigma2 = 1e-5)
    ys[,i] <- rho * ys[,i-1] + e_i
  }
  
  u <- mvtnorm::rmvnorm(t, sigma = sigmaU) 
  d <- t(ds) + u %*% t(A) + rnorm(t * N, sd = 0.1)
  y <- t(ys) + d %*% t(B_true) + u %*% t(gamma) + rnorm(t * N, sd = 0.6)
  
  # fit DML model
  x <- data.frame(s1=rep(s[, 1], each = t), s2=rep(s[, 2], each = t),
                  time = rep(1:t, N))
  # Polynomial spatial terms
  x$s1_sq  <- x$s1^2
  x$s2_sq  <- x$s2^2
  x$s1_s2  <- x$s1 * x$s2
  x$s1_cu  <- x$s1^3
  x$s2_cu  <- x$s2^3

  # Temporal polynomials
  x$time_sq <- x$time^2
  x$time_cu <- x$time^3

  # Interaction between space and time
  x$s1_time <- x$s1 * x$time
  x$s2_time <- x$s2 * x$time
  x$s1_s2_time <- x$s1 * x$s2 * x$time

  # Periodic time terms (useful if there's any seasonality)
  x$sin_time <- sin(2 * pi * x$time / t)
  x$cos_time <- cos(2 * pi * x$time / t)
  dml_rf <- dml_plm(as.matrix(x), as.vector(d), as.vector(y), dreg, yreg, nfold=5)
  
  # Loop over models
  for (model in c("single_dml", "multi_dml", "stack_dml", "dml_mfactor",
                  "nc_mfactor", "nc_2mfactor",
                  "ife", "dml_ife", "proximal_dnc")) {
    estimate <- fit_spatial_invariant(model, s, d, y, m, dml_rf)  # Fit the model
    print(model)
    results_st_loc[[length(results_st_loc) + 1]] <- data.frame(
      simulation = j,
      model = model,
      estimate = estimate
    )
  }
}

# Combine results into a data frame
results2_df <- do.call(rbind, results_st_loc)
results2_df$error <- (results2_df$estimate - 1) / 1
results2_df <- results2_df %>%
  mutate(model = case_when(
    model == "single_dml" ~ "Single DML",
    model == "multi_dml" ~ "Multi DML",
    model == "stack_dml" ~ "Stacked DML",
    model == "nc_mfactor" ~ "FC(3)",
    model == "nc_2mfactor" ~ "FC(6)",
    model == "dml_mfactor" ~ "FC(3)+DML",
    model == "ife" ~ "IFE(6)",
    model == "dml_ife" ~ "IFE(3)+DML",
    model == "proximal_dnc" ~ "PCI",
    TRUE ~ model
  ))

saveRDS(results2_df, file = "sim_linear_factor_N50T100_st.rds")
```

```{r}
# calculate bias for each model
bias_df <- results2_df %>%
  group_by(model) %>%
  summarize(bias = mean(error))

# calculate sd
sd_df <- results2_df %>%
  group_by(model) %>%
  summarize(sd = sd(estimate))

# Error for each setting

## results2_df <- readRDS("data/sim_linear_factor_N500T10_sp.rds")
p2 <- ggplot(results2_df |> filter(model != "PCI"), 
  aes(x = model, y = error)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red", size = 0.5) +
  geom_boxplot(fill = "dark gray") +
  # facet_wrap(~setting, scales = "free") +
  labs(title = "Spatiotemporal Effects and Spatially Invariant Confounding", 
       x = "Model", 
       y = "Relative Error",
       fill = "Model") +
  theme_bw(base_size=16) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
p2

# ggsave("sim_linear_factor_N50T100_random_error.pdf", p1, width = 6, height = 5, units = "in", dpi = 300)
```

## Time-invariant confounding

```{r}
fit_time_invariant <- function(model, s, d, y, m, cache = NULL) {
  n <- nrow(s)
  t <- ncol(d)

  switch(model,
         single_dml = {
           dml_rf <- dml_plm(s, d[,1], y[,1], dreg, yreg, nfold=5)
           return(dml_rf$coef.est)
         },

         multi_dml = {
           resY <- matrix(cache$resY, nrow=n)
           resD <- matrix(cache$resD, nrow=n)
           lm_model <- lm(resY ~ resD)
           return(coef(lm_model)[2, 1])
         },
         
         stack_dml = {
           lm_model <- lm(cache$resY ~ cache$resD)
           return(coef(lm_model)[2])
         },

         nc_mfactor = {
           sm_nc_factor <- cmdstanr::cmdstan_model("stan/multi_nc_linear_factor.stan")
           data_list <- list(K=t, Q=t, N=n, M=m, d=d, y=y)
           opt_result <- sm_nc_factor$optimize(data=data_list, init=2, iter=1000, show_messages=FALSE)
           return(opt_result$mle("beta"))
         },

         nc_2mfactor = {
           sm_nc_factor <- cmdstanr::cmdstan_model("stan/multi_nc_linear_factor.stan")
           data_list <- list(K=t, Q=t, N=n, M=m*2, d=d, y=y)
           opt_result <- sm_nc_factor$optimize(data=data_list, init=2, iter=1000, show_messages=FALSE)
           return(opt_result$mle("beta"))
         },

         dml_mfactor = {
           resY <- matrix(cache$resY, nrow=n)
           resD <- matrix(cache$resD, nrow=n)
           sm_nc_factor <- cmdstanr::cmdstan_model("stan/multi_nc_linear_factor.stan")
           data_list <- list(K=t, Q=t, N=n, M=m, d=resD, y=resY)
           opt_result <- sm_nc_factor$optimize(data=data_list, init=2, iter=1000, show_messages=FALSE)
           return(opt_result$mle("beta"))
         },
         
         ife = {
           df <- data.frame(unit = rep(1:n, t),
                            time = rep(1:t, each = n),
                            Y = c(y),
                            D = c(d))
           gs <- interFE(Y ~ D, data = df, index = c("unit","time"), r = m*2,
                         force = "none", se = FALSE)
           return(gs$beta)
         },
         
         dml_ife = {
           df <- data.frame(unit = rep(1:n, t),
                            time = rep(1:t, each = n),
                            Y = c(cache$resY),
                            D = c(cache$resD))
           gs <- interFE(Y ~ D, data = df, index = c("unit","time"), r = m,
                         force = "none", se = FALSE)
           return(gs$beta)
         },
         
         proximal_dnc = {                 
           resY <- matrix(cache$resY, nrow = n)   # T × n
           resD <- matrix(cache$resD, nrow = n)
        
           if (n < m + 1)
             stop("Need at least m+1 units to supply ", m, " NC pairs")
  
           Y <- c(resY[, 2:(t - 1)])      # previous outcome and next exposure as NC
           A <- c(resD[, 2:(t - 1)])
           W <- c(resY[, 1:(t - 2)])
           Z <- c(resD[, 3:t])

           p2sls_result <- p2sls.lm(
             Y = Y,
             A = A,
             X = NULL,
             W = W,
             Z = Z,
             nco_type = "linear",
             nco_args = list(list(offset = rep(0, length(Y))))
           )
           return(p2sls_result$ESTIMATE[2])
         },

         NA)
}
```

### Spatial variation

```{r}
set.seed(2025) 
t <- 10 # number of time points or 
n <- 500 # number of locations
m <- 3 # numer of latent confounders
sigmaU <- diag(m) 
results_sp_time <- list()
n_simulations <- 100

A <- matrix(rnorm(t * m, mean = 0, sd=0.3), nrow = t, ncol = m)
R <- qr.Q(qr(matrix(rnorm(m*m), m))) 
gamma <- 0.2 * (A %*% R) + 0.5 * A
B_true <- diag(t)
  
for (i in 1:n_simulations) {
  # Generate data for the current setting
  s <- matrix(runif(2 * n, -1, 1), ncol = 2) # N by 2
  fs <- sin(2*pi*s[,1]*s[,2]) + s[,1] + s[,2] # N by 1
  u <- mvtnorm::rmvnorm(n, sigma = sigmaU) # t * m
  d <- u %*% t(A) + matrix(rep(fs^2/2, t), nrow=n, byrow=FALSE) + rnorm(n * t, sd = 0.2) # n by t
  y <- d %*% t(B_true) + u %*% t(gamma) + matrix(rep(fs*2, t), nrow=n, byrow=FALSE) + rnorm(n * t, sd = 0.5) 
  
  # fit DML model
  x <- data.frame(s1=rep(s[, 1], t), s2=rep(s[, 2], t))
  # Polynomial spatial terms
  x$s1_sq  <- x$s1^2
  x$s2_sq  <- x$s2^2
  x$s1_s2  <- x$s1 * x$s2
  x$s1_cu  <- x$s1^3
  x$s2_cu  <- x$s2^3
  dml_rf <- dml_plm(as.matrix(x), as.vector(d), as.vector(y), dreg, yreg, nfold=5)

  # Loop over models
  for (model in c("single_dml", "multi_dml", "stack_dml", "dml_mfactor",
                  "nc_mfactor", "nc_2mfactor",
                  "ife", "dml_ife", "proximal_dnc")) {
    estimate <- fit_time_invariant(model, s, d, y, m, dml_rf)  # Fit the model
    print(model)
    results_sp_time[[length(results_sp_time) + 1]] <- data.frame(
      simulation = i,
      model = model,
      estimate = estimate
    )
  }
}
results3_df <- do.call(rbind, results_sp_time)
results3_df$error <- (results3_df$estimate - 1) / 1 
results3_df <- results3_df %>%
  mutate(model = case_when(
    model == "single_dml" ~ "Single DML",
    model == "multi_dml" ~ "Multi DML",
    model == "stack_dml" ~ "Stacked DML",
    model == "nc_mfactor" ~ "FC(3)",
    model == "nc_2mfactor" ~ "FC(6)",
    model == "dml_mfactor" ~ "FC(3)+DML",
    model == "ife" ~ "IFE(6)",
    model == "dml_ife" ~ "IFE(3)+DML",
    model == "proximal_dnc" ~ "PCI",
    TRUE ~ model
  ))

# saveRDS(results3_df, file = "sim_linear_factor_N500T10_sp.rds")
```

```{r}
## results3_df <- readRDS("data/sim_linear_factor_N50T100_st.rds")
p3 <- ggplot(results3_df |> filter(model != "PCI"), aes(x = model, y = error)) +
  geom_boxplot(fill = "dark gray") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red", size = 0.5) +
  # facet_wrap(~setting, scales = "free") +
  labs(title = "Spatiotemporal Effects and Time-Invariant Confounding", 
       x = "Model", 
       y = "Relative Error",
       fill = "Model") +
  theme_bw(base_size=16) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") + 
  ylim(-0.2, 0.5)
p3

# ggsave("sim_linear_factor_N500T10_sp_error.pdf", p3, width = 6, height = 5, units = "in", dpi = 300)
```

### Spatiotemporal variation

```{r}
set.seed(2025) 
t <- 10  # number of time points
n <- 500  # number of locations
m <- 3  # number of latent confounders
sigmaU <- diag(m)  
results_st_time <- list()
rho <- 0.9
n_simulations <- 100

A <- matrix(rnorm(t * m, mean = 0, sd=0.5), nrow = t, ncol = m)
R <- qr.Q(qr(matrix(rnorm(m*m), m))) 
gamma <- 0.2 * (A %*% R) + 0.5 * A

for (j in 1:n_simulations) {
  # Generate spatio-temporal component
  s <- matrix(runif(2 * n, -1, 1), ncol = 2) # n by 2 locations
  ds <- matrix(0, nrow=n, ncol=t) 
  ds[,1] <- gp_sample(s, a2 = 1, l = 0.5, sigma2 = 1e-5)
  for(i in 2:t) {
    e_i <- gp_sample(s, a2 = 0.1, l = 0.5, sigma2 = 1e-5)
    ds[,i] <- rho * ds[,i-1] + e_i
  }
  ys <- matrix(0, nrow=n, ncol=t) # n by t 
  ys[,1] <- gp_sample(s, a2 = 1, l = 0.5, sigma2 = 1e-5)
  for(i in 2:t) {
    e_i <- gp_sample(s, a2 = 0.1, l = 0.5, sigma2 = 1e-5)
    ys[,i] <- rho * ys[,i-1] + e_i
  }
  
  # Generate treatment and outcome
  u <- mvtnorm::rmvnorm(n, sigma = sigmaU) # n * m
  d <- ds + u %*% t(A) + rnorm(n * t, sd = 0.2) # n by t
  B_true <- diag(t)
  y <- d %*% t(B_true) + ys + u %*% t(gamma) + rnorm(n * t, sd = 0.5)
           
  # fit DML model
  x <- data.frame(s1=rep(s[, 1], t), s2=rep(s[, 2], t), time = rep(1:t, each = n))
  # Polynomial spatial terms
  x$s1_sq  <- x$s1^2
  x$s2_sq  <- x$s2^2
  x$s1_s2  <- x$s1 * x$s2
  
  # Temporal polynomials
  x$time_sq <- x$time^2
  
  # Interaction between space and time
  x$s1_time <- x$s1 * x$time
  x$s2_time <- x$s2 * x$time
  x$s1_s2_time <- x$s1 * x$s2 * x$time

  # Periodic time terms (useful if there's any seasonality)
  x$sin_time <- sin(2 * pi * x$time / t)
  x$cos_time <- cos(2 * pi * x$time / t)
  dml_rf <- dml_plm(as.matrix(x), as.vector(d), as.vector(y), dreg, yreg, nfold=5)
  
  # Loop over models
  for (model in c("single_dml", "multi_dml", "stack_dml", "dml_mfactor",
                  "nc_mfactor", "nc_2mfactor",
                  "ife", "dml_ife", "proximal_dnc")) {
    estimate <- fit_time_invariant(model, s, d, y, m, dml_rf)  # Fit the model
    print(model)
    results_st_time[[length(results_st_time) + 1]] <- data.frame(
      simulation = j,
      model = model,
      n = n,
      t = t,
      estimate = estimate
    )
  }
}
results4_df <- do.call(rbind, results_st_time)
results4_df$error <- results4_df$estimate - 1
results4_df <- results4_df %>%
  mutate(model = case_when(
    model == "single_dml" ~ "Single DML",
    model == "multi_dml" ~ "Multi DML",
    model == "stack_dml" ~ "Stacked DML",
    model == "nc_mfactor" ~ "FC(3)",
    model == "nc_2mfactor" ~ "FC(6)",
    model == "dml_mfactor" ~ "FC(3)+DML",
    model == "ife" ~ "IFE(6)",
    model == "dml_ife" ~ "IFE(3)+DML",
    model == "proximal_dnc" ~ "PCI",
    TRUE ~ model
  ))

# saveRDS(results4_df, file = "sim_linear_factor_N500T10_st.rds")
```

```{r}
## results4_df <- readRDS("data/sim_linear_factor_N500T10_st.rds")
p4 <- ggplot(results4_df |> filter(model != "PCI"), aes(x = model, y = error)) +
  geom_boxplot(fill = "dark gray") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red", size = 0.5) +
  # facet_wrap(~setting, scales = "free") +
  labs(title = "Spatiotemporal Effects and Time-Invariant Confounding", 
       x = "Model", 
       y = "Relative Error",
       fill = "Model") +
  theme_bw(base_size=16) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") + 
  ylim(-0.2, 0.6)
p4

# ggsave("sim_linear_factor_N500T10_st_error.pdf", p4, width = 6, height = 5, units = "in", dpi = 300)
```


## Different adjustment sets

```{r}
fit_adj_model <- function(model, s, d, y, m, cache = NULL) {
  n <- nrow(s)
  t <- ncol(d)
  time <- 1:t

  switch(model,
         stack_dml = {
           lm_model <- lm(cache$resY ~ cache$resD)
           return(coef(lm_model)[2])
         },

         dml_factor = {
           resY <- matrix(cache$resY, nrow=n)
           resD <- matrix(cache$resD, nrow=n)
           sm_nc_factor <- cmdstanr::cmdstan_model("stan/multi_nc_linear_factor.stan")
           data_list <- list(K=t, Q=t, N=n, M=m, d=resD, y=resY)
           opt_result <- sm_nc_factor$optimize(data=data_list, init=2, iter=1000, show_messages=FALSE)
           return(opt_result$mle("beta"))
         },
         
         dml_ife = {
           df <- data.frame(unit = rep(1:n, t),
                            time = rep(1:t, each = n),
                            Y = c(cache$resY),
                            D = c(cache$resD))
           gs <- interFE(Y ~ D, data = df, index = c("unit","time"), r = m,
                         force = "none", se = FALSE)
           return(gs$beta)
         },

         NA)
}
```

```{r}
N = 500
t = 10
beta = 1
rho_H = 0.9
rho_V = 0.8   

set.seed(2025)
results_adjust <- list()
n_simulations <- 1

a_ST <- 0.8; a_U1 <- rnorm(t, 0.5, 0.1); # a_U2 <- rnorm(t, -0.3, 0.1);
a_V <- 0.4; a_S <- 0.4; a_W <- 0.4
g_ST <- 0.6; g_U1 <- rnorm(t, 0.5, 0.1); # g_U2 <- rnorm(t, -0.2, 0.1);
g_V <- 0.3; g_S <- 0.4; g_W <- 0.4

  
for (j in 1:n_simulations) {
  s <- matrix(runif(N * 2, -1, 1), ncol = 2)
  distM  <- rdist(s)
  K <- exp(- distM^2 / (2 * 0.4^2)) + diag(1e-6, N)
  
  ## spatial GP for ST interaction term
  G_i <- mvrnorm(1, mu = rep(0, N), Sigma = K)
  
  ## second spatial GP 
  S_i <- mvrnorm(1, mu = rep(0, N), Sigma = K)
  
  ## time-invariant non-spatial confounder
  U_i1 <- rnorm(N, sd = 1)
  # U_i2 <- rnorm(N, sd = 1)
  
  ## AR-1 temporal driver for ST term
  H_t <- numeric(t)
  H_t[1] <- rnorm(1, mean = 1, sd = 0.3)
  for(i in 2:t) H_t[i] <- rho_H * H_t[i-1] + rnorm(1, sd = sqrt(1 - rho_H^2))
  
  ## AR-1 time-varying confounder
  V_t <- numeric(t)
  V_t[1] <- rnorm(1, mean = 1, sd = 0.3)
  for(i in 2:t) V_t[i] <- rho_V * V_t[i-1] + rnorm(1, sd = sqrt(1 - rho_V^2))
  
  ## spatiotemporal signal
  ST_it <- outer(G_i, H_t)               # N × T
  
  W_it <- matrix(rnorm(N*t, 0, 1), N, t) 
  # W_it <- outer(rnorm(N, 1, 0.5), arima.sim(list(ar = 0.9), n = t))
  # w_i <- matrix(rnorm(N*5), N, 5)
  # w_t <- matrix(arima.sim(list(ar = 0.8), n = t*5), t, 5)
  # W_it <- w_i %*% t(w_t)
  
  # Generate D and Y
  eps_D <- matrix(rnorm(N*t, 0, 0.3), N, t)   # measurement noise
  eps_Y <- matrix(rnorm(N*t, 0, 0.5), N, t)
  d <- a_ST*ST_it + # spatiotemporal component
       outer(U_i1, a_U1) +   # static (non-spatial) confounder
       # outer(a_V, V_t) + # time-varying confounder
       a_V*matrix(rep(V_t, each=N), N) +
       a_W*W_it + # time and spatial-varying
       eps_D # varying coefficient?
  
  y <- beta*d +
       g_ST*ST_it +
       g_S*S_i +
       outer(U_i1, g_U1) + 
       # outer(g_V, V_t) +
       g_V*matrix(rep(V_t, each=N), N) +
       g_W*W_it +
       eps_Y
  
  # # fit DML model
  x <- data.frame(w = as.vector(W_it))
  # s1=rep(s[, 1], t), s2=rep(s[, 2], t), time = rep(1:t, each = N) , u = rep(U_i, t), w= W_it)
  
  # # Polynomial spatial terms
  # x$s1_sq  <- x$s1^2
  # x$s2_sq  <- x$s2^2
  # x$s1_s2  <- x$s1 * x$s2
  # #
  # # # Temporal polynomials
  # x$time_sq <- x$time^2
  # #
  # # # Interaction between space and time
  # x$s1_time <- x$s1 * x$time
  # x$s2_time <- x$s2 * x$time
  # x$s1_s2_time <- x$s1 * x$s2 * x$time
  # #
  # # # Periodic time terms (useful if there's any seasonality)
  # x$sin_time <- sin(2 * pi * x$time / t)
  # x$cos_time <- cos(2 * pi * x$time / t)
  # 
  dml_rf <- dml_plm(NULL, as.vector(d), as.vector(y), dreg, yreg, nfold=5)
  
  # Loop over models
  for (model in c("stack_dml", "dml_factor","dml_ife")) {
    estimate <- fit_adj_model(model, s, d, y, 3, dml_rf)  # Fit the model
    print(model)
    results_adjust[[length(results_adjust) + 1]] <- data.frame(
      simulation = j,
      model = model,
      adj = "None",
      estimate = estimate
    )
  }
}

results_adj_df <- do.call(rbind, results_adjust)
results_adj_df$error <- results_adj_df$estimate - 1
results_adj_df
```

```{r}
results_adj_df <- results_adj_df %>%
  mutate(adj = case_when(
    adj == "No Adjustment" ~ "None",
    adj == "All Confounders" ~ "All",
    TRUE ~ adj
  ))

results_adj_df <- results_adj_df %>%
  mutate(model = case_when(
    model == "stack_dml" ~ "Naive",
    model == "dml_factor" ~ "FC",
    model == "dml_ife" ~ "IFE",
    TRUE ~ model
  ))

results_adj_df$adj <- factor(results_adj_df$adj,
  levels = c(
    "None",
    "Unit-specific",
    "Spatiotemporal",
    "All")
)
results_adj_df$model <- factor(results_adj_df$model, levels = c("Naive", "IFE", "FC"))

p5 <- ggplot(results_adj_df,
       aes(x     = adj,        # groups on the x-axis
           y     = estimate,
           fill  = model)) +              # colour = model

  geom_boxplot(position = position_dodge(width = 0.7), width = 0.6) +
  geom_hline(yintercept = 1,
             linetype   = "dashed",
             colour     = "red",
             linewidth  = 0.4) +

  labs(title = "Estimated Causal Effect by Model and Adjustment Set",
       x     = "Adjustment Set",
       y     = "Estimated Effect",
       fill  = "Model") +

  scale_fill_brewer(palette = "Set2") +
  coord_cartesian(ylim = c(1, 2)) +
  theme_minimal(base_size = 11) +
  theme(axis.text.x  = element_text(angle = 45, hjust = 1))
p5
# ggsave("sim_adjustment_N500T10.pdf", p5, width = 5, height = 5, units = "in", dpi = 300)
```

## Nonlinear effect

```{r}
# Simulation setting
set.seed(123)
n  <- 1000      # sample size
d  <- 5         # # treatments / outcomes
M  <- 2         # # latent factors
p  <- 2         # # observed confounders

# Coefficients
beta_X  <- matrix(c( 1.0, -1.2,   # for D1
                     0.8,  0.5,   # for D2
                    -0.4,  1.0,   
                     0.6, -0.7,
                     1.5,  0.9), nrow = d, byrow = TRUE)   # d × p

A_mat      <- matrix(c(1.0,  0.3,
                       0.4, -0.8,
                      -0.6,  1.0,
                      -0.7,  -0.3,
                       1.0, -0.5), nrow = d, byrow = TRUE) 
Gamma_mat  <- matrix(c(0.4, -0.7,
                      -0.3,  0.2,
                       0.8,  0.2,
                       0.2,  0.7,
                       0.5,  0.4), nrow = d, byrow = TRUE) 

# Generate data
X        <- matrix(rnorm(n * p), n, p)              # (n × 2) observed confounders
colnames(X) <- c("X1", "X2")

V        <- matrix(rnorm(n * M), n, M)              # (n × 2) latent factors
sigma_xi_vec  <- runif(d, 0.3, 0.5)  
sigma_eps_vec <- runif(d, 0.3, 0.5)

xi <- sapply(sigma_xi_vec, function(sd) rnorm(n, sd = sd))
eps <- sapply(sigma_eps_vec, function(sd) rnorm(n, sd = sd))

fD <- function(x1, x2, beta) {        # non‑linear f_Dj
  beta[1] * sin(x1) + beta[2] * (x2^2 - 1)   # zero‑mean in x2
}

D <- matrix(NA, n, d)
for (j in 1:d) {
  mu_D      <- fD(X[,1], X[,2], beta_X[j, ])
  D[, j]    <- mu_D + V %*% A_mat[j, ] + xi[, j]
}
colnames(D) <- paste0("D", 1:d)

g_fun_list <- list(
  function(d, x1, x2) tanh(d) + 0.8 * x1^2 - 0.6 * exp(-x2),
  function(d, x1, x2) sin(d) + 0.7 * x1 * x2,
  function(d, x1, x2) - 0.3 * d^2 + 0.5 * x1 + 0.6 * x2^2,
  function(d, x1, x2) d^3 / (1 + d^2) + 0.3 * x1^2 - 0.4 * x2,
  function(d, x1, x2) log(abs(d) + 1) + 0.4 * x1 - 0.2 * x2^2
)

Y <- matrix(NA, n, d)
for (j in 1:d) {
  g_j      <- g_fun_list[[j]]
  mu_Y     <- g_j(D[, j], X[, 1], X[, 2])
  Y[, j]   <- mu_Y + V %*% Gamma_mat[j, ] + eps[, j]
}
colnames(Y) <- paste0("Y", 1:d)

sim_data <- data.frame(X, D, Y)
head(sim_data)

# Bias
Sigma_D  <- A_mat %*% t(A_mat) + diag(sigma_xi_vec^2)
Sigma_VD <- diag(M) - t(A_mat) %*% solve(Sigma_D) %*% A_mat
Sigma_VD_invhalf <- expm::sqrtm(solve(Sigma_VD))
tilde_Gamma <- Gamma_mat %*% Sigma_VD_invhalf   #  Γ Σ_{V|D,X}^{-1/2}
C_true <- Gamma_mat %*% Sigma_VD_invhalf %*% t(A_mat) %*%
          solve(Sigma_D)  
ind_bias <- diag(diag(d) %*% tilde_Gamma %*% t(A_mat) %*% diag(d) / Sigma_D)
print(ind_bias)
```

Estimate bias matrix.

```{r}
set.seed(123)

# -------------------------------------------
# 0.  Helper: n‑fold cross‑fitting indices
# K_fold <- 5
# fold_id <- sample(rep(1:K_fold, length.out = nrow(sim_data)))

# -------------------------------------------
# 1. Estimate off-diagonal of C
# for (j in 1:d) {
#   # storage for residuals
#   residD <- matrix(NA, nrow(sim_data), d-1)
#   colnames(residD) <- paste0("D", setdiff(1:d, j))
#   residY <- numeric(nrow(sim_data))
# 
#   for (fold in 1:K_fold) {
#     train <- fold_id != fold
#     test  <- fold_id == fold
# 
#     # flexible model for each D_{-j} and Y_j (here: thin‑plate splines)
#     for (k in setdiff(1:d, j)) {
#       form_k <- as.formula(
#         paste0("D", k, " ~ ",
#                "s(X1, k = 10) + s(X2, k = 10) + ",
#                "s(D", j, ", k = 10)")
#       )
#       fDk <- gam(form_k, data = sim_data[train, ])
#       residD[test, paste0("D", k)] <- sim_data[test, paste0("D", k)] -
#         predict(fDk, newdata = sim_data[test, ])
#     }
#     form_Yj <- as.formula(
#       paste0("Y", j, " ~ ",
#              "s(X1, k = 10) + s(X2, k = 10) + ",
#              "s(D", j, ", k = 10)")
#     )
#     fYj <- gam(form_Yj, data = sim_data[train, ])
#     residY[test] <-
#       sim_data[test, paste0("Y", j)] -
#       predict(fYj, newdata = sim_data[test, ])
#   }
#   fit_j <- lm(residY ~ residD - 1)
#   C_hat[j, -j] <- coef(fit_j)
# }

# for (fold in 1:K_fold) {
#   train <- fold_id != fold;  test <- !train
#   

d <- 5  # number of treatments / outcomes
M <- 2  # number of latent factors we simulated
C_hat <- matrix(0, d, d)

for (j in 1:d) {
  form <- as.formula(
    paste0("Y",j, " ~ ",
           "s(D",j,", k = 10) +",          # flexible g_j(·)
           "s(X1, k = 10) + s(X2, k = 10) + ",
           paste(paste0("D", setdiff(1:d, j)), collapse = " + "))
  )
  fit <- mgcv::gam(form, data = sim_data)
  
  # linear coefficients come last in the param vector
  C_hat[j, -j] <- coef(fit)[paste0("D", setdiff(1:d, j)) ]
}

off_diag_mask <- row(C_hat) != col(C_hat)
print(sum(abs((C_hat-C_true)[off_diag_mask])))
C_off <- C_hat  # store for Procrustes

# -------------------------------------------
# 2. Factor model on residuals and pinpoint C
D_tilde <- matrix(NA, nrow(sim_data), d)
colnames(D_tilde) <- paste0("D", 1:d)
for (j in 1:d) {
    fDj <- gam(sim_data[,paste0("D",j)] ~ s(X1)+s(X2),
               data=sim_data)
    D_tilde[,j] <- sim_data[,paste0("D",j)] - predict(fDj, sim_data)
}

Y_tilde <- matrix(NA, nrow(sim_data), d)
colnames(Y_tilde) <- paste0("Y", 1:d)
for (j in 1:d) {
  form_Yj <- as.formula(
    paste0("Y", j, " ~ ",
           "s(X1, k = 10) + s(X2, k = 10) + ",
           "s(D", j, ", k = 10) + ",
           paste(paste0("D", setdiff(1:d, j)), collapse = " + "))
    )
  fYj <- gam(form_Yj, data = sim_data)
  Y_tilde[,j] <- sim_data[,paste0("Y",j)] - predict(fYj, sim_data)
}
fa_res <- factanal(D_tilde, factors = M)
A_hat <- as.matrix(fa_res$loadings)     # d × M
sds   <- apply(D_tilde, 2, sd)       # standard deviations of columns
A_hat <- A_hat * sds                 # scale loadings back to original scale
psi_raw <- fa_res$uniquenesses         # vector of length d
Psi_cov <- diag(psi_raw * sds^2)        # scale to original variance scale
Sigma_D_hat <- A_hat %*% t(A_hat) + Psi_cov  

fa_resY <- factanal(Y_tilde, M) # fa(D_tilde, nfactors = M, rotate = "none", scores = "regression")
Gamma_hat <- as.matrix(fa_resY$loadings)   # d × M
Gamma_hat <- Gamma_hat * sqrt(diag(cov(Y_tilde)))         

Sigma_VD_hat <- diag(M) - t(A_hat) %*% solve(cov(D_tilde)) %*% A_hat
Sigma_VD_hat <- (Sigma_VD_hat + t(Sigma_VD_hat))/2   # numerical stabilisation
Sigma_VD_hat <- Sigma_VD_hat + 1e-8 * diag(M)

inv_sqrt_Sigma_VD <- expm::sqrtm(solve(Sigma_VD_hat))
R_hat  <- inv_sqrt_Sigma_VD %*% t(A_hat) %*% solve(cov(D_tilde))

# Optimize on the Stiefel manifold
# -------- objective:  ‖  W*(Γ Θ R − C_off) ‖_F²  --------------------
Wmat <- matrix(1, d, d); diag(Wmat) <- 0
obj_fn <- function(Theta) {
  Mtmp  <- Gamma_hat %*% Theta %*% R_hat          # d × d
  diff  <- (Mtmp - C_off) * Wmat                  # mask diag
  sum(diff * diff)                                # Frobenius²
}

# gradient 
grad_fn <- function(Theta) {
  Mtmp  <- Gamma_hat %*% Theta %*% R_hat
  diff  <- (Mtmp - C_off) * Wmat                  # d × d
  2 * ( t(Gamma_hat) %*% diff %*% t(R_hat) )      # M × M
}

set.seed(111)
best_obj <- Inf
Theta_hat <- NULL
n_starts <- 10  # try 5-20 for stability
for (i in 1:n_starts) {
  Theta_init <- rstiefel::rustiefel(M, M) 
  current_result <- rstiefel::optStiefel(F = obj_fn,
                          dF = grad_fn,
                          Vinit = Theta_init, 
                          # method="curvilinear",
                          maxIters = 100,
                          maxLineSearchIters = 100,
                          tol = 1e-10,
                          verbose = TRUE) # M × M orthogonal
  current_obj <- obj_fn(current_result)
  if (current_obj < best_obj) {
    best_obj <- current_obj
    Theta_hat <- current_result
  }
}

# --------- recovered bias matrix ------------------------------------
C_hat <- Gamma_hat %*% Theta_hat %*% R_hat
C_hat1 <- C_off
diag(C_hat1) <- diag(C_hat)
print(round(C_true,3))
print(round(C_hat1, 3))
print(round(C_hat, 3))
print(round(max(abs(C_hat - C_true)),3))

# -------------------------------------------
# 3.  Debias Y and estimate g_j(d)
Y_db <- as.matrix(sim_data[, paste0("Y", 1:d)]) -
        D_tilde %*% t(C_hat1)
# Y_db <- as.matrix(sim_data[, paste0("Y", 1:d)]) -
#         D_tilde %*% diag(ind_bias)

# Dose–response at mean X  (we ignore X to mimic "average over X")
x1_bar <- mean(sim_data$X1)
x2_bar <- mean(sim_data$X2)

g_fit  <- list()
d_grid <- seq(-5, 5, length = 300)
g_hat  <- matrix(NA, length(d_grid), d)

for (j in 1:d) {
  df <- data.frame(y = Y_db[, j],
                   d = sim_data[, paste0("D", j)],
                   x1 = sim_data$X1,
                   x2 = sim_data$X2)
  g_fit[[j]] <- gam(y ~ s(d, k = 20) + s(x1, k = 20) + s(x2, k = 20), data = df)
  g_hat[, j] <- predict(g_fit[[j]], newdata = data.frame(
    d = d_grid,
    x1 = x1_bar,   # Fix x1, x2 at mean so only s(d) varies
    x2 = x2_bar
  ))   #,  type = "terms")[, "s(d)"]
}
plot_df <- data.frame(
  d = rep(d_grid, d),
  g = as.vector(g_hat),
  outcome = factor(rep(1:d, each = length(d_grid)), labels = paste("Outcome", 1:d))
)


# -------------------------------------------
# 4.  Plot of the five causal curves

# Create true g_j curves at mean(X1, X2)
g_true <- matrix(NA, length(d_grid), d)
for (j in 1:d) {
  g_j <- g_fun_list[[j]]
  g_true[, j] <- g_j(d_grid, x1_bar, x2_bar)
}

# Prepare data frame for true g_j
true_df <- data.frame(
  d = rep(d_grid, d),
  g = as.vector(g_true),
  outcome = factor(rep(1:d, each = length(d_grid)), labels = paste("Outcome", 1:d))
)

# Create naive estimate 
g_fit_naive  <- list()
g_hat_naive  <- matrix(NA, length(d_grid), d)

for (j in 1:d) {
  df_naive <- data.frame(y = sim_data[, paste0("Y", j)],
                         d = sim_data[, paste0("D", j)],
                         x1 = sim_data$X1,
                         x2 = sim_data$X2)
  g_fit_naive[[j]] <- gam(y ~ s(d, k = 20) + s(x1, k = 20) + s(x2, k = 20), data = df_naive)
  g_hat_naive[,j] <- predict(g_fit_naive[[j]], newdata = data.frame(
    d = d_grid,
    x1 = x1_bar,   # Fix x1, x2 at mean so only s(d) varies
    x2 = x2_bar
  ))         # ,  type = "terms")[, "s(d)"]
}
plot_df_naive <- data.frame(
  d = rep(d_grid, d),
  g = as.vector(g_hat_naive),
  outcome = factor(rep(1:d, each = length(d_grid)), labels = paste("Outcome", 1:d))
)

# plot

p6 <- ggplot() +
  geom_line(data = plot_df, aes(x = d, y = g, color = "Debiased"), size = 1) +
  geom_line(data = plot_df_naive, aes(x = d, y = g, color = "Naive"), linetype = "dotdash", size = 1) +
  geom_line(data = true_df, aes(x = d, y = g, color = "True"), linetype = "dashed", size = 1) +
  facet_wrap(~ outcome, scales = "free_y", ncol = 3) +
  scale_y_continuous(breaks = pretty_breaks(n = 5)) +
  scale_color_manual(
    values = c(
      "Debiased" = "#0072B2",  # brighter blue
      "Naive" = "#009E73",     # brighter green
      "True" = "#E41A1C"       # brighter red-orange
    ),
    labels = c("Debiased Estimate", "Naive Estimate", "True Effect")
  ) +
  labs(
    title = "Estimated Dose–Response Functions",
    x = expression("Difference from Dose 0"),
    y = expression("Causal Effect"),
    color = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 12),
    strip.text = element_text(face = "bold", size = 12),
    panel.grid.major = element_line(size = 0.2, color = "grey80"),
    panel.grid.minor = element_blank(),
    text = element_text(family = "serif"),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
p6
# ggsave("sim_nonlinear_Factor_5outcomes.pdf", p6, width = 6, height = 5, units = "in", dpi = 300)
```

Another estimator

```{r}
# Factor loadings for exposures
D_tilde <- matrix(NA, nrow(sim_data), d)
colnames(D_tilde) <- paste0("D", 1:d)
for (j in 1:d) {
    fDj <- gam(sim_data[,paste0("D",j)] ~ s(X1)+s(X2),
               data=sim_data)
    D_tilde[,j] <- resid(fDj)
}
fa_res <- factanal(D_tilde, factors = M)
A_hat <- as.matrix(fa_res$loadings)     # d × M
sds   <- apply(D_tilde, 2, sd)       # standard deviations of columns
A_hat <- A_hat * sds                 # scale loadings back to original scale
psi_raw <- fa_res$uniquenesses         # vector of length d
Psi_cov <- diag(psi_raw * sds^2)        # scale to original variance scale
Sigma_D_hat <- A_hat %*% t(A_hat) + Psi_cov  
print(Sigma_D_hat)
```

```{r}
library(splines)
library(glmmTMB)

long <- sim_data |>
  mutate(row_id = row_number()) |>
  pivot_longer(cols = starts_with("Y"),
               names_to  = "resp",
               values_to = "Y") |>
  mutate(resp = factor(resp, levels = paste0("Y",1:d)))

## Dummy matrix for responses
resp_mat <- model.matrix(~ resp - 1, long)   # n×d, columns respY1 … respY5

## Helper to add basis × response interaction
add_basis <- function(var, df = 6) {
  B   <- bs(long[[var]], df = df)
  out <- purrr::imap_dfc(as.data.frame(B), \(col, k){
           base <- paste0("b", var, "_", k)
           # multiply by each response dummy -> own columns
           purrr::imap_dfc(as.data.frame(resp_mat), \(dum, rnm){
               nm <- paste0(base, "_", rnm)
               setNames(col * dum, nm)
           })
         })
  out
}

## Build bases for D1..D5 and X1, X2
basis_D <- purrr::map_dfc(paste0("D",1:d), add_basis, df = 6)
basis_X1 <- add_basis("X1", df = 6)
basis_X2 <- add_basis("X2", df = 6)

long <- bind_cols(long, basis_D, basis_X1, basis_X2)

############################################################
## 4.  BUILD FORMULA  ######################################
############################################################
## (i) Causal matrix: resp:Dk gives one coefficient per Y_j × D_k
CD_terms <- paste0("resp:D", 1:d, collapse = " + ")

## (ii) Spline columns: include everything that starts with 'b'
spline_cols <- grep("^b", names(long), value = TRUE)
spline_terms <- paste(spline_cols, collapse = " + ")

form <- as.formula(
  paste("Y ~ resp +",          # intercepts
        CD_terms, "+",
        spline_terms,
        "+ rr(resp + 0 | row_id, d = 2)"))   # rank-2 residual factors

############################################################
## 5.  FIT  ###############################################
############################################################
fit_Y <- glmmTMB(form, data = long, family = gaussian())

############################################################
## 6.  RESULTS  ############################################
############################################################
cat("\n=== Estimated causal-effect matrix  Ĉ  ====================\n")
C_est <- summary(fit_Y)$coeff$cond |>
  subset(grepl("^respY[1-5]:D", rownames(.))) |>
  transform(Y = sub(":.*","",rownames(.)),
            D = sub(".*:","",rownames(.))) |>
  xtabs(Estimate ~ Y + D, data = _)
print(round(C_est,3))

cat("\n=== Γ̂  (residual factor loadings from rr()) ===============\n")
print(round(getME(fit_Y, "fact_load")[[1]],3))
```

## Identification without factor model

```{r}
N <- 10
t <- 100000
m <- 3
sigmaU <- diag(m)  # Update sigmaU based on N

A <- matrix(rnorm(N * m, mean = 0, sd=0.5), nrow = N, ncol = m)
R <- qr.Q(qr(matrix(rnorm(m*m), m))) 
gamma <- 0.3 * (A %*% R) + 0.6 * A
gamma <- matrix(rnorm(N * m, mean = 0, sd=0.5), nrow = N, ncol = m)
B_true <- diag(N)

Sigma_D  <- A %*% t(A) + diag(N) * 0.5^2
C_true <- gamma %*% t(A) %*% solve(Sigma_D)  
naive_true <- B_true + C_true

# Generate data
# s <- matrix(runif(2 * N, -1, 1), ncol = 2) # N by 2
# fs <- sin(2*pi*s[,1]*s[,2]) + s[,1] + s[,2] # N by 1
u <- mvtnorm::rmvnorm(t, sigma = sigmaU) # t * m

d <- u %*% t(A) + rnorm(t * N, sd = 0.5) # t by N
y <- d %*% t(B_true) + u %*% t(gamma) + rnorm(t * N, sd = 0.5) 

      
naive_hat  <- ((t(y) %*% d) / t) %*% solve(crossprod(d) / t)
print(norm(naive_hat - naive_true, "F"))


f <- function(b){
  sv <- svd(naive_hat - b * diag(N))$d
  sum(tail(sort(sv, decreasing = FALSE), N - m)^2)
}
beta_hat <- optimize(f, interval = c(-10, 10))$minimum
C_hat    <- A_hat - beta_hat * diag(N)

eig_vals <- eigen(naive_mat, only.values = TRUE)$values
eig_vals_sorted <- sort(Re(eig_vals))           # ensure real/ordered

beta_hat <- mean(eig_vals_sorted[1:(N - m)])    # mean of (N-m) repeated eigenvalues
C_hat    <- naive_mat - beta_hat * diag(N)          # subtract β I
print(norm(C_hat - C_true, "F"))
print(beta_hat)
```

## Interference

```{r}
fit_spatial_inter <- function(model, d, d_nb, y, m) {
  t <- nrow(d); n <- ncol(d)

  switch(model,

    # -------------------------------------------------- 1. Naïve OLS
    naive = {
      df <- data.frame(
        Y    = c(y),
        D    = c(d),
        D_nb = c(d_nb)
      )
      ols <- lm(Y ~ D + D_nb, data = df)
      return(coef(ols)[c("D", "D_nb")])        #  (β1, β2)
    },

    # -------------------------------------------------- 2. Factor model
    nc_factor = {
      sm_factor <- cmdstanr::cmdstan_model(
        "stan/spatial_inter_linear_factor.stan"
      )
      data_list <- list(K = n, N = t, M = m,
                        d = d, y = y, d_nb = d_nb) 
      opt_result <- sm_factor$optimize(data = data_list,
                                       init = 2, iter = 1000,
                                       show_messages = FALSE)
      return(opt_result$mle(c("beta1", "beta2")))
    },

    # -------------------------------------------------- 3. IFE / interactive FE
    ife = {
      df <- data.frame(unit = rep(1:n, each = t),
                       time = rep(1:t,  n),
                       Y = c(y),
                       D = c(d),  
                       D_nb = c(d_nb))
      gs <- interFE(
              Y ~ D + D_nb,
              data  = df,
              index = c("unit", "time"),
              r     = m,
              force = "none",
              se    = FALSE
            )
      return(gs$beta)
    },

    stop("Unknown model name")
  )
}
```

```{r}
# rho <- 0.9
models <- c("naive", "nc_factor", "ife")
results_inter_sim <- data.frame() # will grow row-wise

set.seed(2025) 
n_sim = 100

N <- 50
t <- 200
m <- 4
sigmaU <- diag(m) 
A_self <- matrix(rnorm(N * (m-1), mean = 0, sd=0.5), nrow = N, ncol = (m-1))
A_nb_col <- rnorm(N, 0, 0.5)  
A <- cbind(A_self,  A_nb_col)    
R <- qr.Q(qr(matrix(rnorm((m-1) * (m-1)), (m-1)))) 
gamma_self <- 0.3 * (A_self %*% R) + 0.5 * A_self
gamma_nb_col  <- W %*% A_nb_col
gamma  <- cbind(gamma_self, gamma_nb_col)

B_true <- diag(N)

## interference
k_neigh   <- 3          # number of nearest neighbours whose exposure spills over
beta_self <- 1.0        # effect of the unit’s own exposure D_it
beta_nb   <- 0.5        # effect of neighbours’ average exposure

for (sim in seq_len(n_sim)) {
  print(paste("iteration", sim))
  s <- matrix(runif(2 * N, -1, 1), ncol = 2) # N by 2
  
  ## Build a k-NN weight matrix W  (rows = recipients i, cols = senders j)
  knn_idx <- get.knn(s, k = k_neigh)$nn.index   # N × k matrix of neighbour indices
  
  W <- matrix(0, N, N)                          # raw adjacency
  for (i in 1:N) W[i, knn_idx[i, ]] <- 1
  W <- W / rowSums(W)                           # row-standardize to averages
  
  ## Exposure-to-outcome coefficient matrix that embeds interference
  B_true <- beta_self * diag(N) + beta_nb * W   # N × N  (diag = own effect)
  
  ## Simulate factor confounders, exposures D, and outcomes Y
  u <- mvtnorm::rmvnorm(t, sigma = sigmaU)      # t × m   (time rows)
  
  d <- u %*% t(A) + rnorm(t * N, sd = 0.2)      # t(ds) + exposures  (t × N)
  d_nb <- d %*% t(W)                    # t by N matrix of neighbour averages

  y <-  d %*% t(B_true) +                       # own + neighbour exposures
        u %*% t(gamma) +                        # shared factor confounding
        rnorm(t * N, sd = 0.5)                  # idiosyncratic noise
                                                # t(ys) +
  for (mod in models) {
    theta_hat <- fit_spatial_inter(mod, d, d_nb, y, m)

    results_inter_sim <- rbind(
      results_inter_sim,
      data.frame(sim      = sim,
                 model    = mod,
                 beta_self_hat = theta_hat[1],
                 beta_nb_hat   = theta_hat[2])
    )
  }
}
results_inter_sim <- results_inter_sim %>%
  mutate(model = case_when(
    model == "naive" ~ "DML (NUC)",
    model == "nc_factor" ~ "FC",
    model == "ife" ~ "IFE",
    TRUE ~ model
  ))
head(results_inter_sim)
```

```{r}
results_long <- results_inter_sim %>%     
  transmute(model,
            Direct   = beta_self_hat,
            Indirect = beta_nb_hat) %>%                            # keep estimates
  pivot_longer(cols = c(Direct, Indirect),
               names_to   = "effect",
               values_to  = "estimate") %>%
  mutate(true = ifelse(effect == "Direct",  beta_self, beta_nb),
         rel_error = (estimate - true) / true)   

results_long$model <- factor(results_long$model, levels = c("DML (NUC)", "IFE", "FC"))

p7 <- ggplot(results_long,
       aes(x = model, y = estimate)) +
  geom_boxplot(width = 0.7, alpha = 0.8, fill="dark gray", outliers=FALSE) +
  geom_hline(data = distinct(results_long, effect, true),
             aes(yintercept = true), linetype = "dashed",
             colour = "red", size = 0.4) +
  facet_wrap(~ effect, scales = "free_y") +
  labs(title = "Estimated Direct and Nearest Neighbor Effects",
       x = "Estimator",
       y = "Estimated Effect",
       fill = "Estimator") +
  scale_fill_brewer(palette = "Set2") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")

ggsave("sim_interference.png", p7, width = 8, height = 5)
```

```{r}
# saveRDS(results_long, file = "sim_spatial_interference_N50T200.rds")
# ggsave("sim_spatial_interference_N50T200.pdf", p7, width = 6, height = 5, units = "in", dpi = 300)
```

## Violation of normality and linearity 

```{r}
## ============================================================
## Misspecification simulations for FC / IFE / NUC
##   Part 1: Non-Gaussian confounder / noise
##   Part 2: Nonlinear exposure / outcome confounding
## ============================================================

## ------------------------------------------------------------
## Helpers
## ------------------------------------------------------------

# Standardize a matrix columnwise
scale_signal <- function(X) {
  X <- scale(X, center = TRUE, scale = TRUE)
  as.matrix(X)
}

# Center panel only (no variance scaling)
center_panel <- function(X) {
  X <- scale(X, center = TRUE, scale = FALSE)
  as.matrix(X)
}

# rank-r truncated SVD reconstruction
svd_reconstruct <- function(W, r) {
  s <- svd(W)
  if (r == 0) {
    return(matrix(0, nrow = nrow(W), ncol = ncol(W)))
  }
  U_r <- s$u[, 1:r, drop = FALSE]
  V_r <- s$v[, 1:r, drop = FALSE]
  D_r <- diag(s$d[1:r], nrow = r, ncol = r)
  U_r %*% D_r %*% t(V_r)
}

# Bai-Ng-style rank selection on a reduced-form joint panel:
# W = [D, Y - beta_tilde D], where beta_tilde is the pooled OLS slope.
choose_rank_bn_joint <- function(d, y, rank_candidates, ic_type = "IC2") {
  beta_tilde <- coef(lm(as.vector(y) ~ as.vector(d)))[2]
  y_tilde <- y - beta_tilde * d

  W <- cbind(
    scale(d, center = TRUE, scale = FALSE),
    scale(y_tilde, center = TRUE, scale = FALSE)
  )

  TT <- nrow(W)
  PP <- ncol(W)

  max_rank_allowed <- min(TT, PP) - 1L
  rank_candidates <- sort(unique(rank_candidates))
  rank_candidates <- rank_candidates[rank_candidates <= max_rank_allowed]
  if (length(rank_candidates) == 0) {
    stop("No candidate ranks are allowed by the panel dimensions.")
  }

  ic_vals <- numeric(length(rank_candidates))
  names(ic_vals) <- rank_candidates

  for (j in seq_along(rank_candidates)) {
    r <- rank_candidates[j]
    W_hat <- svd_reconstruct(W, r)
    V_r <- mean((W - W_hat)^2)

    penalty <- switch(
      ic_type,
      IC1 = r * ((TT + PP) / (TT * PP)) * log((TT * PP) / (TT + PP)),
      IC2 = r * ((TT + PP) / (TT * PP)) * log(min(TT, PP)),
      IC3 = r * (log(min(TT, PP)) / min(TT, PP)),
      stop("Unknown ic_type")
    )

    ic_vals[j] <- log(V_r) + penalty
  }

  as.integer(rank_candidates[which.min(ic_vals)])
}

# Error generator
gen_error <- function(type, Tt, N, sigma_target,
                      df_t_err = 4,
                      hetero_scale = NULL,
                      skew_alpha = 5) {

  if (type == "gaussian") {
    e <- matrix(rnorm(Tt * N, 0, 1), Tt, N)

  } else if (type == "t_df") {
    e <- matrix(rt(Tt * N, df = df_t_err), Tt, N)

  } else if (type == "uniform") {
    # Uniform(-sqrt(3), sqrt(3)) has Var = 1
    e <- matrix(runif(Tt * N, min = -sqrt(3), max = sqrt(3)), Tt, N)

  } else if (type == "skew_norm") {
    if (!requireNamespace("sn", quietly = TRUE)) {
      stop("skew_norm requested but package 'sn' is not installed.")
    }
    e <- matrix(sn::rsn(Tt * N, xi = 0, omega = 1, alpha = skew_alpha), Tt, N)

  } else if (type == "hetero") {
    if (is.null(hetero_scale)) stop("hetero_scale required for hetero")
    raw <- matrix(rt(Tt * N, df = df_t_err), Tt, N)
    e <- raw * matrix(hetero_scale, Tt, N, byrow = TRUE)

  } else {
    stop("Unknown error type: ", type)
  }

  e <- e - mean(e)
  e <- e / sd(as.vector(e)) * sigma_target
  e
}

# Discrete latent factor generator
gen_discrete_U <- function(t, m, K = 5, prob = NULL, jitter_sd = 0) {
  if (is.null(prob)) prob <- rep(1 / K, K)

  U_class <- matrix(rnorm(K * m), nrow = K, ncol = m)
  k_t <- sample(1:K, t, replace = TRUE, prob = prob)
  u <- U_class[k_t, , drop = FALSE]

  if (jitter_sd > 0) {
    u <- u + matrix(rnorm(t * m, mean = 0, sd = jitter_sd), nrow = t, ncol = m)
  }

  u <- scale(u, center = TRUE, scale = TRUE)
  as.matrix(u)
}

# Nonlinear transform applied to linear indices, not directly to U
apply_nonlinearity <- function(X, type) {
  out <- switch(
    type,
    cubic = X^3,
    linear_plus_cubic = X + 0.3 * X^3,
    tanh = tanh(X),
    sine = sin(X),
    stop("Unknown nonlinearity type: ", type)
  )
  scale_signal(out)
}

# Run estimators
run_estimators <- function(s, d, y, m_fit, dreg, yreg,
                           benchmark_rank = 3,
                           rank_candidates = c(1, 2, 3, 4, 5, 6, 8, 10),
                           max_ident_rank = max(rank_candidates),
                           rank_ic_type = "IC2",
                           sm_nc_factor = NULL) {
  d <- center_panel(d)
  y <- center_panel(y)

  dml_rf <- NULL
  invisible(capture.output({
    dml_rf <- dml_plm(NULL, as.vector(d), as.vector(y), dreg, yreg, nfold = 5)
  }))
  dml_rf$sm_nc_factor <- sm_nc_factor

  fc_rank <- if (identical(m_fit, "selected")) {
    choose_rank_bn_joint(
      d = d,
      y = y,
      rank_candidates = rank_candidates,
      ic_type = rank_ic_type
    )
  } else {
    as.integer(m_fit)
  }

  enriched_rank <- min(2L * fc_rank, as.integer(max_ident_rank))
  benchmark_rank <- as.integer(benchmark_rank)

  out <- list()
  for (model in c("stack_dml", "nc_mfactor", "nc_mfactor_enriched", "ife")) {
    model_rank <- if (model == "stack_dml") NA_integer_ else benchmark_rank
    if (model == "nc_mfactor") model_rank <- fc_rank
    if (model == "nc_mfactor_enriched") model_rank <- enriched_rank
    fit_model <- if (model == "nc_mfactor_enriched") "nc_mfactor" else model
    est <- fit_spatial_invariant(fit_model, s, d, y, model_rank, dml_rf)
    out[[length(out) + 1]] <- data.frame(
      model = model,
      estimate = unname(est),
      fitted_rank = model_rank,
      rank_rule = dplyr::case_when(
        model == "nc_mfactor" && identical(m_fit, "selected") ~ paste0("Bai-Ng ", rank_ic_type),
        model == "nc_mfactor_enriched" && identical(m_fit, "selected") ~ paste0("min(2 * Bai-Ng ", rank_ic_type, ", Mmax)"),
        TRUE ~ "fixed"
      )
    )
  }
  bind_rows(out)
}

# Label models
label_models <- function(df) {
  df %>%
    mutate(
      model = case_when(
        model == "stack_dml" ~ "NUC",
        model == "nc_mfactor" ~ ifelse(rank_rule == "fixed", "FC", "FC(selected)"),
        model == "nc_mfactor_enriched" ~ "FC(enriched)",
        model == "ife" ~ "IFE",
        TRUE ~ model
      )
    ) %>%
    mutate(model = factor(model, levels = c("NUC", "IFE", "FC", "FC(selected)", "FC(enriched)")))
}

## ------------------------------------------------------------
## Shared parameter generator
## ------------------------------------------------------------

make_params <- function(N, m) {
  A <- matrix(rnorm(N * m, mean = 0, sd = 0.5), nrow = N, ncol = m)
  R <- qr.Q(qr(matrix(rnorm(m * m), m)))
  gamma <- 0.3 * (A %*% R) + 0.5 * A
  B_true <- diag(N)

  list(A = A, gamma = gamma, B_true = B_true)
}

## ============================================================
## Part 1: Non-Gaussian confounder / noise
## ============================================================

run_nonGaussian_sim <- function(
    n_simulations = 50,
    N = 50,
    t = 100,
    m = 3,
    sigma_D_err = 0.7,
    sigma_Y_err = 0.7,
    df_t_err = 4,
    skew_alpha = 5,
    discrete_prob = c(0.1, 0.15, 0.2, 0.25, 0.3),
    m_fit = m,
    benchmark_rank = m,
    rank_candidates = c(1, 2, 3, 4, 5, 6, 8, 10),
    max_ident_rank = max(rank_candidates),
    rank_ic_type = "IC2",
    dreg,
    yreg,
    seed = 2025) {

  set.seed(seed)

  # Dropped mixture; keep uniform + discrete
  dist_types <- c("gaussian", "t_df", "uniform", "skew_norm", "hetero", "discreteU")

  params <- make_params(N, m)
  A <- params$A
  gamma <- params$gamma
  B_true <- params$B_true
  sigmaU <- diag(m)
  sm_nc_factor <- cmdstanr::cmdstan_model("stan/multi_nc_linear_factor.stan")

  results <- list()

  for (i in seq_len(n_simulations)) {
    cat("nonGaussian iteration", i, "\n")

    s <- matrix(runif(2 * N, -1, 1), ncol = 2)
    fs <- sin(2 * pi * s[, 1] * s[, 2]) + s[, 1] + s[, 2]
    hetero_scale_D <- 1 + 0.5 * abs(fs)
    hetero_scale_Y <- 1 + 0.5 * abs(fs)

    for (ftype in dist_types) {

      if (ftype == "discreteU") {
        u <- gen_discrete_U(t = t, m = m, K = 5, prob = discrete_prob, jitter_sd = 0)
        epsD <- gen_error("gaussian", t, N, sigma_D_err, df_t_err = df_t_err)
        epsY <- gen_error("gaussian", t, N, sigma_Y_err, df_t_err = df_t_err)

      } else {
        if (ftype == "hetero") {
          u <- mvtnorm::rmvnorm(t, sigma = sigmaU)
        } else {
          u <- gen_error(
            type = ftype,
            Tt = t,
            N = m,
            sigma_target = 1,
            df_t_err = df_t_err,
            skew_alpha = skew_alpha
          )
        }

        epsD <- gen_error(
          type = ftype,
          Tt = t,
          N = N,
          sigma_target = sigma_D_err,
          df_t_err = df_t_err,
          hetero_scale = if (ftype == "hetero") hetero_scale_D else NULL,
          skew_alpha = skew_alpha
        )

        epsY <- gen_error(
          type = ftype,
          Tt = t,
          N = N,
          sigma_target = sigma_Y_err,
          df_t_err = df_t_err,
          hetero_scale = if (ftype == "hetero") hetero_scale_Y else NULL,
          skew_alpha = skew_alpha
        )
      }

      d <- u %*% t(A) + epsD
      y <- d %*% t(B_true) + u %*% t(gamma) + epsY

      fit_df <- run_estimators(
        s, d, y,
        m_fit = m_fit,
        benchmark_rank = benchmark_rank,
        dreg = dreg,
        yreg = yreg,
        rank_candidates = rank_candidates,
        max_ident_rank = max_ident_rank,
        rank_ic_type = rank_ic_type,
        sm_nc_factor = sm_nc_factor
      )
      fit_df$simulation <- i
      fit_df$facet <- ftype

      results[[length(results) + 1]] <- fit_df
    }
  }

  bind_rows(results) %>%
    label_models() %>%
    mutate(
      facet = case_when(
        facet == "gaussian"  ~ "Gaussian",
        facet == "t_df"      ~ "Student-t",
        facet == "uniform"   ~ "Uniform",
        facet == "skew_norm" ~ "Skewed",
        facet == "hetero"    ~ "Heteroskedastic",
        facet == "discreteU" ~ "Discrete",
        TRUE ~ facet
      )
    ) %>%
    mutate(
      facet = factor(
        facet,
        levels = c("Gaussian", "Student-t", "Uniform",
                   "Skewed", "Heteroskedastic", "Discrete")
      )
    )
}

## ============================================================
## Part 2: Nonlinear exposure / nonlinear outcome confounding
## ============================================================

run_nonlinear_sim <- function(
    n_simulations = 50,
    N = 50,
    t = 100,
    m = 3,
    sigma_D_err = 0.7,
    sigma_Y_err = 0.7,
    exposure_types = c("linear_plus_cubic", "cubic", "tanh"),
    outcome_types = c("linear_plus_cubic", "sine", "tanh"),
    both_pairs = list(),
    m_fit = m,
    benchmark_rank = m,
    rank_candidates = c(1, 2, 3, 4, 5, 6, 8, 10),
    max_ident_rank = max(rank_candidates),
    rank_ic_type = "IC2",
    dreg,
    yreg,
    seed = 2026,
    exposure_loading_scale = 1,
    outcome_loading_scale = 1,
    outcome_nonlinear_scale = 1) {

  set.seed(seed)

  params <- make_params(N, m)
  A <- params$A
  gamma <- params$gamma
  B_true <- params$B_true
  sigmaU <- diag(m)
  sm_nc_factor <- cmdstanr::cmdstan_model("stan/multi_nc_linear_factor.stan")
  
  nice_label <- function(x) {
    dplyr::case_when(
      x == "linear_plus_cubic" ~ "Lin + Cubic",
      x == "cubic" ~ "Cubic",
      x == "tanh" ~ "Tanh",
      x == "sine" ~ "Sine",
      TRUE ~ x
    )
  }

  results <- list()

  for (i in seq_len(n_simulations)) {
    cat("nonlinear iteration", i, "\n")

    s <- matrix(runif(2 * N, -1, 1), ncol = 2)

    u <- mvtnorm::rmvnorm(t, sigma = sigmaU)
    epsD <- gen_error("gaussian", t, N, sigma_D_err)
    epsY <- gen_error("gaussian", t, N, sigma_Y_err)
    
    A_weak <- exposure_loading_scale * A
    gamma_strong <- outcome_loading_scale * gamma

    lin_D <- u %*% t(A_weak)
    lin_Y <- u %*% t(gamma_strong)

    # Nonlinear exposure settings
    for (etype in exposure_types) {
      d_signal <- apply_nonlinearity(lin_D, etype)
      d <- d_signal + epsD
      y <- d %*% t(B_true) + lin_Y + epsY

      fit_df <- run_estimators(
        s, d, y,
        m_fit = m_fit,
        benchmark_rank = benchmark_rank,
        dreg = dreg,
        yreg = yreg,
        rank_candidates = rank_candidates,
        max_ident_rank = max_ident_rank,
        rank_ic_type = rank_ic_type,
        sm_nc_factor = sm_nc_factor
      )
      fit_df$simulation <- i
      fit_df$facet <- paste0("Exp: ", nice_label(etype))
      results[[length(results) + 1]] <- fit_df
    }

    # Nonlinear outcome-confounding settings
    for (ytype in outcome_types) {
      conf_signal <- outcome_nonlinear_scale * apply_nonlinearity(lin_Y, ytype)
      d <- lin_D + epsD
      y <- d %*% t(B_true) + conf_signal + epsY

      fit_df <- run_estimators(
        s, d, y,
        m_fit = m_fit,
        benchmark_rank = benchmark_rank,
        dreg = dreg,
        yreg = yreg,
        rank_candidates = rank_candidates,
        max_ident_rank = max_ident_rank,
        rank_ic_type = rank_ic_type,
        sm_nc_factor = sm_nc_factor
      )
      fit_df$simulation <- i
      fit_df$facet <- paste0("Out: ", nice_label(ytype))
      results[[length(results) + 1]] <- fit_df
    }
    
    # both nonlinear
    for (pair in both_pairs) {
      etype <- pair[1]
      ytype <- pair[2]

      d_signal <- apply_nonlinearity(lin_D, etype)
      conf_signal <- outcome_nonlinear_scale * apply_nonlinearity(lin_Y, ytype)

      d <- d_signal + epsD
      y <- d %*% t(B_true) + conf_signal + epsY

      fit_df <- run_estimators(
        s, d, y,
        m_fit = m_fit,
        benchmark_rank = benchmark_rank,
        dreg = dreg,
        yreg = yreg,
        rank_candidates = rank_candidates,
        max_ident_rank = max_ident_rank,
        rank_ic_type = rank_ic_type,
        sm_nc_factor = sm_nc_factor
      )
      fit_df$simulation <- i
      fit_df$facet <- paste0("Both: ", nice_label(etype), " / ", nice_label(ytype))
      results[[length(results) + 1]] <- fit_df
    }
  }

  dplyr::bind_rows(results) %>%
    label_models()
}

## ============================================================
## Plotting helpers
## ============================================================

plot_nonGaussian <- function(df) {
  ggplot(df, aes(x = model, y = estimate, fill = model)) +
    geom_hline(yintercept = 1, linetype = "dashed", color = "red", linewidth = 0.4) +
    geom_boxplot(fill = "dark gray") +
    facet_wrap(~ facet, nrow = 2, scales = "free_y") +
    labs(
      title = "Estimates under Non-Gaussian Confounder and Noise Distributions",
      x = NULL,
      y = "Estimated effect"
    ) +
    theme_bw() +
    theme(
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.6),
      strip.background = element_rect(fill = "grey90", color = "black", linewidth = 0.6),
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none",
      plot.title = element_text(face = "bold"),
      plot.margin = ggplot2::margin(8, 8, 8, 8)
    )
}

plot_nonlinear <- function(df) {
  ggplot(df, aes(x = model, y = estimate, fill = model)) +
    geom_hline(yintercept = 1, linetype = "dashed", color = "red", linewidth = 0.4) +
    geom_boxplot(fill = "dark gray") +
    facet_wrap(~ facet, nrow = 2, scales = "free_y") +
    labs(
      title = "Estimates under Nonlinear Misspecification",
      x = NULL,
      y = "Estimated effect"
    ) +
    theme_bw() +
    theme(
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.6),
      strip.background = element_rect(fill = "grey90", color = "black", linewidth = 0.6),
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none",
      plot.title = element_text(face = "bold"),
      plot.margin = ggplot2::margin(8, 8, 8, 8)
    )
}

plot_nonlinear <- function(df, nrow = 2) {
  facet_levels <- unique(df$facet)
  
  facet_labs <- c(
    "Exp: Lin + Cubic" = "Exp: Lin + Cubic",
    "Exp: Cubic" = "Exp: Cubic",
    "Exp: Tanh" = "Exp: Tanh",
    "Out: Lin + Cubic" = "Out: Lin + Cubic",
    "Out: Sine" = "Out: Sine",
    "Out: Tanh" = "Out: Tanh",
    "Both: Lin + Cubic / Lin + Cubic" = "Both:\nLin + Cubic / Lin + Cubic",
    "Both: Cubic / Sine" = "Both:\nCubic / Sine",
    "Both: Tanh / Tanh" = "Both:\nTanh / Tanh"
  )

  ggplot(df, aes(x = model, y = estimate, fill = model)) +
    geom_hline(yintercept = 1, linetype = "dashed", color = "red", linewidth = 0.4) +
    geom_boxplot(fill = "dark gray") +
    facet_wrap(
      ~ facet,
      nrow = nrow,
      scales = "free_y",
      labeller = as_labeller(facet_labs)
    ) +
    labs(
      title = "Estimates under Nonlinear Misspecification",
      x = NULL,
      y = "Estimated effect"
    ) +
    theme_bw() +
    theme(
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.6),
      strip.background = element_rect(fill = "grey90", color = "black", linewidth = 0.6),
      strip.text = element_text(size = 10, face = "bold"),
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none",
      plot.title = element_text(face = "bold"),
      plot.margin = ggplot2::margin(8, 8, 8, 8)
    )
}

```

```{r}
## ============================================================
## Run simulations
## ============================================================

# You must already have:
#   dml_plm(...)
#   fit_spatial_invariant(...)
#   dreg
#   yreg
# defined in your environment.

# Part 1: Non-Gaussian
results_nonGaussian_df <- run_nonGaussian_sim(
  n_simulations = 50,
  N = 30,
  t = 100,
  m = 3,
  sigma_D_err = 0.7,
  sigma_Y_err = 0.7,
  m_fit = "selected",
  dreg = dreg,
  yreg = yreg,
  seed = 2025
)

# Part 2: Nonlinear
results_nonlinear_df <- run_nonlinear_sim(
  n_simulations = 50,
  N = 30,
  t = 100,
  m = 3,
  sigma_D_err = 0.7,
  sigma_Y_err = 0.7,
  exposure_types = c("linear_plus_cubic", "cubic", "tanh"),
  outcome_types = c("linear_plus_cubic", "sine", "tanh"),
  m_fit = "selected",
  dreg = dreg,
  yreg = yreg,
  seed = 2026,
  exposure_loading_scale = 1,
  outcome_loading_scale = 1,
  outcome_nonlinear_scale = 1
)

saveRDS(results_nonGaussian_df, file = "data/sim_nonGaussian_N30T100.rds")
saveRDS(results_nonlinear_df, file = "data/sim_nonlinear_N30T100.rds")

## ============================================================
## Summaries
## ============================================================

summary_nonGaussian <- results_nonGaussian_df %>%
  group_by(facet, model) %>%
  summarise(
    mean_est = mean(estimate),
    bias = mean(estimate - 1),
    sd_est = sd(estimate),
    rmse = sqrt(mean((estimate - 1)^2)),
    .groups = "drop"
  )

summary_nonlinear <- results_nonlinear_df %>%
  group_by(facet, model) %>%
  summarise(
    mean_est = mean(estimate),
    bias = mean(estimate - 1),
    sd_est = sd(estimate),
    rmse = sqrt(mean((estimate - 1)^2)),
    .groups = "drop"
  )

rank_nonGaussian <- results_nonGaussian_df %>%
  filter(model %in% c("FC(selected)", "FC(enriched)")) %>%
  count(facet, model, fitted_rank, name = "n") %>%
  group_by(facet, model) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

rank_nonlinear <- results_nonlinear_df %>%
  filter(model %in% c("FC(selected)", "FC(enriched)")) %>%
  count(facet, model, fitted_rank, name = "n") %>%
  group_by(facet, model) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

print(summary_nonGaussian)
print(summary_nonlinear)
print(rank_nonGaussian)
print(rank_nonlinear)

## ============================================================
## Figures
## ============================================================

p_nonGaussian <- plot_nonGaussian(results_nonGaussian_df)
p_nonlinear  <- plot_nonlinear(results_nonlinear_df)

print(p_nonGaussian)
print(p_nonlinear)

saveRDS(rank_nonGaussian, file = "data/sim_nonGaussian_rank_sel_N30T100.rds")
saveRDS(rank_nonlinear, file = "data/sim_nonlinear_rank_sel_N30T100.rds")

dir.create("figures", showWarnings = FALSE)

ggsave("figures/sim_nonGaussian_N30T100.pdf", p_nonGaussian,
       width = 8, height = 5.5, units = "in", dpi = 300)

ggsave("figures/sim_nonlinear_N30T100.pdf", p_nonlinear,
       width = 8, height = 5.5, units = "in", dpi = 300)
```


## Rank comparison

```{r}
## ============================================================
## Exact factor rank sensitivity under spatially invariant confounding
## Rank selection via Bai-Ng IC on [D, Y - beta_tilde D]
## Compare FC(3), FC(6), FC(selected)
## ============================================================

set.seed(2026)

# -----------------------------
# Settings
# -----------------------------
n_simulations <- 50
N <- 30
t <- 100
beta0 <- 1

true_rank_grid <- c(3, 6, 10)
fit_rank_grid  <- c(3, 6)

# candidate ranks for selected FC
rank_candidates <- c(1, 2, 3, 4, 5, 6, 8, 10, 12, 15)

sigma_D_err <- 0.5
sigma_Y_err <- 0.5

# -----------------------------
# Helpers
# -----------------------------
scale_panel <- function(M) {
  M <- scale(M, center = TRUE, scale = FALSE)
  as.matrix(M / sd(as.vector(M)))
}

# rank-r truncated SVD reconstruction
svd_reconstruct <- function(W, r) {
  s <- svd(W)
  if (r == 0) {
    return(matrix(0, nrow = nrow(W), ncol = ncol(W)))
  }
  U_r <- s$u[, 1:r, drop = FALSE]
  V_r <- s$v[, 1:r, drop = FALSE]
  D_r <- diag(s$d[1:r], nrow = r, ncol = r)
  U_r %*% D_r %*% t(V_r)
}

# Bai-Ng style IC on reduced-form joint panel:
# W = [D, Y - beta_tilde D]
choose_rank_bn_joint <- function(d, y, rank_candidates, ic_type = "IC2") {
  # pooled slope to partially remove direct treatment effect
  beta_tilde <- coef(lm(as.vector(y) ~ as.vector(d)))[2]

  y_tilde <- y - beta_tilde * d

  # center columns
  W <- cbind(
    scale(d, center = TRUE, scale = FALSE),
    scale(y_tilde, center = TRUE, scale = FALSE)
  )

  TT <- nrow(W)
  PP <- ncol(W)

  max_rank_allowed <- min(TT, PP) - 1L
  rank_candidates <- sort(unique(rank_candidates))
  rank_candidates <- rank_candidates[rank_candidates <= max_rank_allowed]

  ic_vals <- numeric(length(rank_candidates))
  names(ic_vals) <- rank_candidates

  for (j in seq_along(rank_candidates)) {
    r <- rank_candidates[j]
    W_hat <- svd_reconstruct(W, r)
    V_r <- mean((W - W_hat)^2)

    penalty <- switch(
      ic_type,
      IC1 = r * ((TT + PP) / (TT * PP)) * log((TT * PP) / (TT + PP)),
      IC2 = r * ((TT + PP) / (TT * PP)) * log(min(TT, PP)),
      IC3 = r * (log(min(TT, PP)) / min(TT, PP)),
      stop("Unknown ic_type")
    )

    ic_vals[j] <- log(V_r) + penalty
  }

  as.integer(rank_candidates[which.min(ic_vals)])
}

# compile once
sm_nc_factor_rank <- cmdstanr::cmdstan_model("stan/multi_nc_linear_factor.stan")

fit_fc_rank_spatial <- function(d, y, M_use) {
  data_list <- list(
    K = ncol(d),   # N
    Q = ncol(y),   # N
    N = nrow(d),   # t
    M = as.integer(M_use),
    d = d,
    y = y
  )

  opt_result <- sm_nc_factor_rank$optimize(
    data = data_list,
    init = 2,
    iter = 1000,
    show_messages = FALSE
  )

  as.numeric(opt_result$mle("beta"))
}

# -----------------------------
# Simulation
# -----------------------------
results_rank_fc <- list()
selected_rank_log <- list()

for (m_true in true_rank_grid) {
  cat("true rank =", m_true, "\n")

  # fixed loadings within each true-rank setting
  A <- matrix(rnorm(N * m_true, mean = 0, sd = 0.5), nrow = N, ncol = m_true)
  R <- qr.Q(qr(matrix(rnorm(m_true * m_true), m_true)))
  gamma <- 0.3 * (A %*% R) + 0.5 * A

  for (i in 1:n_simulations) {

    # latent factors vary over time: t x m_true
    U <- mvtnorm::rmvnorm(t, sigma = diag(m_true))

    # exact shared factor confounding
    conf_D <- scale_panel(U %*% t(A))
    conf_Y <- scale_panel(U %*% t(gamma))

    epsD <- matrix(rnorm(t * N, sd = sigma_D_err), nrow = t, ncol = N)
    epsY <- matrix(rnorm(t * N, sd = sigma_Y_err), nrow = t, ncol = N)

    d <- conf_D + epsD
    y <- beta0 * d + conf_Y + epsY

    d <- scale(d, center = TRUE, scale = FALSE)
    y <- scale(y, center = TRUE, scale = FALSE)

    # better selected rank
    m_hat <- choose_rank_bn_joint(
      d = d,
      y = y,
      rank_candidates = rank_candidates,
      ic_type = "IC2"
    )

    selected_rank_log[[length(selected_rank_log) + 1]] <- data.frame(
      simulation = i,
      true_rank = m_true,
      selected_rank = m_hat
    )

    # fixed-rank fits
    for (m_fit in fit_rank_grid) {
      est <- fit_fc_rank_spatial(d, y, M_use = m_fit)

      results_rank_fc[[length(results_rank_fc) + 1]] <- data.frame(
        simulation = i,
        true_rank = m_true,
        fitted_rank = m_fit,
        model = paste0("FC(", m_fit, ")"),
        estimate = est,
        error = est - beta0
      )
    }

    # selected-rank fit
    est_sel <- fit_fc_rank_spatial(d, y, M_use = m_hat)

    results_rank_fc[[length(results_rank_fc) + 1]] <- data.frame(
      simulation = i,
      true_rank = m_true,
      fitted_rank = m_hat,
      model = "FC(selected)",
      estimate = est_sel,
      error = est_sel - beta0
    )
  }
}

# -----------------------------
# Data frames
# -----------------------------
results_rank_fc_df <- dplyr::bind_rows(results_rank_fc) %>%
  dplyr::mutate(
    model = factor(model, levels = c("FC(3)", "FC(6)", "FC(selected)")),
    true_rank_label = factor(
      paste0("true rank = ", true_rank),
      levels = paste0("true rank = ", true_rank_grid)
    )
  )

selected_rank_df <- dplyr::bind_rows(selected_rank_log) %>%
  dplyr::mutate(
    true_rank_label = factor(
      paste0("true rank = ", true_rank),
      levels = paste0("true rank = ", true_rank_grid)
    )
  )

# -----------------------------
# Summaries
# -----------------------------
summary_rank_fc <- results_rank_fc_df %>%
  dplyr::group_by(true_rank, model) %>%
  dplyr::summarise(
    mean_est = mean(estimate),
    bias = mean(error),
    rmse = sqrt(mean(error^2)),
    sd_est = sd(estimate),
    .groups = "drop"
  )

print(summary_rank_fc)

selected_rank_tab <- selected_rank_df %>%
  dplyr::count(true_rank, selected_rank) %>%
  dplyr::group_by(true_rank) %>%
  dplyr::mutate(prop = n / sum(n)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    true_rank_label = factor(
      paste0("true rank = ", true_rank),
      levels = paste0("true rank = ", true_rank_grid)
    )
  )

print(selected_rank_tab)

saveRDS(results_rank_fc_df, file = "data/sim_rank_fc_N30T100.rds")
saveRDS(selected_rank_tab, file = "data/sim_rank_sel_N30T100.rds")

# -----------------------------
# Main figure: estimation error by true rank
# -----------------------------
p_rank_fc <- ggplot(results_rank_fc_df,
                    aes(x = model, y = error, fill = model)) +
  geom_boxplot(fill = "dark gray", outlier.alpha = 0.3, width = 0.75) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  facet_wrap(~ true_rank_label, nrow = 1, scales = "free_y") +
  labs(
    x = NULL,
    y = expression(hat(beta) - beta),
    title = "FC Sensitivity to Factor-Rank Specification"
  ) +
  theme_bw(base_size = 14) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 25, hjust = 1)
  )

print(p_rank_fc)
dir.create("figures", showWarnings = FALSE)
ggsave("figures/sim_rank_fc_N30T100.pdf", p_rank_fc, width = 8, height = 5, units = "in", dpi = 300)

# -----------------------------
# Supplementary figure: selected rank frequencies
# -----------------------------
p_rank_sel <- ggplot(selected_rank_tab,
                     aes(x = factor(selected_rank), y = prop)) +
  geom_col(width = 0.75, fill = "gray60") +
  facet_wrap(~ true_rank_label, nrow = 1) +
  labs(
    x = "Selected rank",
    y = "Proportion",
    title = "Empirical Distribution of Selected Ranks"
  ) +
  theme_bw(base_size = 14)

print(p_rank_sel)
ggsave("figures/sim_rank_sel_N30T100.pdf", p_rank_sel, width = 8, height = 5, units = "in", dpi = 300)
```


## Exchange spatial and time invariant confounding

```{r}
results_exchange <- list()

set.seed(2025) 
n_simulations = 50

N <- 500
t <- 10
m <- 3
sigmaU <- diag(m) 
A <- matrix(rnorm(N * m, mean = 0, sd=0.5), nrow = N, ncol = m)
R <- qr.Q(qr(matrix(rnorm(m*m), m))) 
gamma <- 0.3 * (A %*% R) + 0.5 * A
B_true <- diag(N)

for (j in 1:n_simulations) {
  print(paste("iteration", j))

  u <- mvtnorm::rmvnorm(t, sigma = sigmaU) 
  d <- u %*% t(A) + rnorm(t * N, sd = 0.5)
  y <- d %*% t(B_true) + u %*% t(gamma) + rnorm(t * N, sd = 1)

  dml_rf <- dml_plm(NULL, as.vector(d), as.vector(y), dreg, yreg, nfold=5)
  
  # Loop over models
  for (model in c("stack_dml", "nc_mfactor", "ife")) {
    estimate <- fit_spatial_invariant(model, s, d, y, m, dml_rf)  # Fit the model
    results_exchange[[length(results_exchange) + 1]] <- data.frame(
      simulation = j,
      model = model,
      estimate = estimate
    )
  }
  model = "nc_mfactor"
  estimate <- fit_time_invariant(model, s, t(d), t(y), m, dml_rf)  # Fit the model
  results_exchange[[length(results_exchange) + 1]] <- data.frame(
    simulation = j,
    model = paste('time', model),
    estimate = estimate
  )
}

# Combine results into a data frame
results_exchange_df <- do.call(rbind, results_exchange)
results_exchange_df <- results_exchange_df %>%
  mutate(model = case_when(
    model == "stack_dml" ~ "NUC",
    model == "nc_mfactor" ~ "Spatial FC",
    model == "ife" ~ "IFE",
    model == "time nc_mfactor" ~ "Time FC",
    TRUE ~ model
  ))

# saveRDS(results_exchange_df, file = "sim_linear_factor_spatial_vs_time.rds")
```

```{r}
results_exchange_df$model <- factor(results_exchange_df$model, levels = c("NUC", "IFE", "Spatial FC", "Time FC"))

p9 <- ggplot(results_exchange_df,
       aes(x = model, y = estimate, fill = model)) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red", size = 0.4) +
  geom_boxplot(outlier.size = 0.7, width = 0.65) +
  labs(title = "Spatially Invariant vs. Time-Invariant Confounding",
       x = "Estimator",
       y = "Estimated Effect",
       fill = "Estimator") +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "grey90", color = NA),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none",
    plot.title = element_text(face = "bold")
  ) +
  coord_cartesian(ylim = c(0.9, 1.5))

# ggsave("sim_spatial_vs_time_N500T10.pdf", p9, width = 6, height = 5, units = "in", dpi = 300)
```

## IFE vs FC under varying individual confounding

```{r}
draw_A_Gamma <- function(N, m, rho, sd_A = 1, sd_Gam = 1) {
  Sigma2 <- matrix(c(1, rho, rho, 1), 2, 2)
  pairs <- rmvnorm(N*m, mean = c(0, 0), sigma = Sigma2)
  A     <- matrix(sd_A  * pairs[, 1], nrow = N, ncol = m)
  Gamma <- matrix(sd_Gam* pairs[, 2], nrow = N, ncol = m)
  list(A = A, Gamma = Gamma)
}
```

```{r}
set.seed(2025)
results_individual <- list()
n_simulations = 20

N <- 50
T_grid  <- c(10, 50, 100)
rho_grid <- c(0.0, 0.5, 0.9)
m <- 3
sigmaU <- diag(m)  


# A <- matrix(rnorm(N * m, mean = 0, sd=0.5), nrow = N, ncol = m)
# R <- qr.Q(qr(matrix(rnorm(m*m), m)))
# gamma <- 0.3 * (A %*% R) + 0.5 * A
beta_true <- 1

for (t in T_grid) {
  print(paste("T=", t))
  for (rho in rho_grid) {
    LG <- draw_A_Gamma(N, m, rho, sd_A = 0.7, sd_Gam = 0.3)
    A <- LG$A; gamma <- LG$Gamma
    # print(cor(as.numeric(A), as.numeric(gamma)))

    for (i in 1:n_simulations) {
      # Generate data
      u <- mvtnorm::rmvnorm(t, sigma = sigmaU) # t * m
      d <- u %*% t(A) + rnorm(t * N, sd = 0.5)  # t by N
      y <- beta_true * d + u %*% t(gamma) + rnorm(t * N, sd = 0.5) 
      
      dml_rf <- dml_plm(NULL, as.vector(d), as.vector(y), dreg, yreg, nfold=5)
    
      # Loop over models
      for (model in c("stack_dml", "nc_mfactor", "ife")) {
        estimate <- fit_spatial_invariant(model, s=NULL, d, y, m, dml_rf)  # Fit the model 
        results_individual[[length(results_individual) + 1]] <- data.frame(
          simulation = i,
          t = t,
          rho = rho,
          model = model,
          estimate = estimate
        )
      }
    }
  }
}


# Combine results into a data frame
results_individual_df <- do.call(rbind, results_individual) %>%
  mutate(model = case_when(
    model == "stack_dml" ~ "NUC",
    model == "nc_mfactor" ~ "FC",
    model == "ife" ~ "IFE",
    TRUE ~ model
  ),
  model = factor(model, levels = c("NUC", "IFE", "FC")),
  rho_f = factor(rho, levels = rho_grid, labels = paste0("rho = ", rho_grid)),
  T_f   = factor(t, levels = T_grid, labels = paste0("T = ", T_grid)))
results_individual_df

results_individual_df <- readRDS("data/sim_IFE_vs_FC_setting_2.rds")
# saveRDS(results_individual_df, file = "data/sim_IFE_vs_FC_setting_1.rds")
```

```{r}
if (!"beta" %in% names(results_individual_df)) results_individual_df$beta <- beta_true

bias_by_Trho <- results_individual_df %>%
  group_by(t, rho, model) %>%
  summarise(
    n     = sum(!is.na(estimate)),
    bias  = mean(estimate - beta, na.rm = TRUE),                 # mean bias
    rmse  = sqrt(mean((estimate - beta)^2, na.rm = TRUE)),       # optional
    .groups = "drop"
  )

print(bias_by_Trho, n=27)

p10 <- ggplot(results_individual_df, aes(x = model, y = estimate)) +
  geom_hline(yintercept = beta_true, linetype = "dashed", color = "red") +
  geom_boxplot(fill = "dark gray", outlier.size = 0.4, width = 0.7) +
  facet_grid(rho_f ~ T_f, scales = "free_y") + 
  labs(x = NULL, y = expression(hat(beta)),
       title = expression("Estimators vs. Confounding Strength and Time Periods")) +
  theme_bw(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 0, hjust = 0.5))
p10

# ggsave("sim_IFE_vs_FC_setting_2.pdf", p10, width = 8, height = 8, units = "in", dpi = 300)
```
